---
title: "Assignment Activity"
author: "VBhagwandas"
date: "24 september 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
First we read the data and then we make the date a date class.

```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date ,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
First we calculate the total number steps in a day:

```{r}
total_steps_day <- aggregate(steps ~ date, activity, sum)
print(total_steps_day)
```

As there are a lot of days, it might be better to represent it in a bar plot:

```{r}
library(ggplot2)
g <- ggplot(total_steps_day, aes(date, steps)) + 
  geom_bar(stat="identity") + 
  labs(x = "Day") + 
  labs(y = "Total steps")

print(g)
```

The median and the mean are as follow:

```{r}
mean_steps <- mean(total_steps_day[,2], na.rm=TRUE)
median_steps <- median(total_steps_day[,2], na.rm=TRUE)
print(mean_steps)
print(median_steps)
```

## What is the average daily activity pattern?
To calculate the average steps taken in the 5 minute interval, represented in a line plot:

```{r}
total_steps_interval <- aggregate(steps ~interval, activity, sum, na.rm=TRUE)
f <- ggplot(total_steps_interval, aes(interval, steps)) + 
  geom_line(stat="identity") + 
  labs(x = "interval") + 
  labs(y = "Total steps")

print(f)
```

The maximun number of stpes in a 5 minute interval across all days is: 

```{r}
max_step_interval <- total_steps_interval[which.max(total_steps_interval[,2]),1]
print(max_step_interval)
```

## Imputing missing values
First we calculate the number of missing values:

```{r}
number_NA <- sum(is.na(activity$steps))
print(number_NA)
```

Now we will impute the missing values with the average of the steps and then we make a histogram:

```{r}
activity$stepsImp <- lapply(activity$steps, function(x)
  if (is.na(x) == TRUE) {activity$stepsImp=round(mean(activity$steps, na.rm=TRUE))}
  else activity$stepsImp=x)

activity$stepsImp <- as.numeric(activity$stepsImp)

total_steps_dayImp <- aggregate(stepsImp ~ date, activity, sum)

h <- ggplot(total_steps_dayImp, aes(date, stepsImp)) + 
  geom_bar(stat="identity") + 
  labs(x = "Day") + 
  labs(y = "Total steps")

print(h)
```

## Are there differences in activity patterns between weekdays and weekends?
First we will assign wether each dat belongs to a weekday or weekend:

```{r}
activity$day <- lapply(activity$date, function(x)
  if (strftime(x, "%u") == 6 | strftime(x, "%u") == 7) 
  {activity$day="weekend"}
  else activity$day="weekday")

activity$day <- as.factor(unlist(activity$day))
```

Then we calculate the steps splitted by interval and weekday/weekend. The results are visualized in a line plot:

```{r}
total_steps_interval2 <- aggregate(steps ~ interval + day, data=activity, sum, na.rm=TRUE)

k <- ggplot(total_steps_interval2, aes(interval, steps)) + 
  geom_line(stat="identity") + 
  facet_grid(.~day) +
  labs(x = "interval") + 
  labs(y = "Total steps")

print(k)
```

Here ends the assignment! 


